### 2016-12-17 2017mergingmerges/section/scrub.tex
ugh

Furthermore, existing systems for combinator fusion only support a subset of combinators, and only in limited cases.
Pull fusion supports joins where one combinator pulls from multiple inputs, but not splits where one output is used my multiple combinators.
Pull can thus express @merge@, but not @partition@.
Push fusion supports splits (@partition@), but not joins (@merge@).

@@

\section{Processes}
\label{s:Process}
Actually this is the process calculus version.

We will start by looking at a simple combinator that increments each value.
Given an input stream @ins@, @increment ins@ creates a new process with one output stream called @outs@.
The process blocks reading the value from @ins@.
After reading the value @i@, it pushes @i + 1@ to the @outs@ stream.
The pulled value must be dropped or released before another value can be performed.
Operationally this is a no-op, but is used for coordinating between producer and consumer.
Finally, the process loops back to the start.

\begin{code}
increment ins = process outs.
  let loop =
     i <- pull ins
     push outs (i + 1)
     drop ins
     jump loop
  in jump loop
\end{code}

@@

\begin{code}
V := Scalar
C := Channel

Process :=
  | V <- pull C
  | push C Exp
  | drop C
  | skip
  | case Exp then Process else Process
  | jump F V*

Program :=
  process C*.
  let F V* = Process
  in  Process
\end{code}

@@

\begin{code}
map f xs = process ys.
  let loop =
     x <- pull xs
     push ys (f x)
     drop xs
     jump loop
  in jump loop
\end{code}

\begin{code}
filter f xs = process ts.
  let loop =
     x <- pull xs
     case f x
       then push ys x
       else skip
     drop xs
     jump loop
  in jump loop
\end{code}

\begin{code}
partition f xs = process ts fs.
  let loop =
     x <- pull xs
     case f x
       then push ts x
       else push fs x
     drop xs
     jump loop
  in jump loop
\end{code}

\begin{code}
zip xs ys = process xys.
  let loop =
     x <- pull xs
     y <- pull ys
     push xys (x,y)
     drop xs
     drop ys
     jump loop
  in jump loop
\end{code}

\begin{code}
merge xs ys = process xys.
  let loop x y =
     case x < y
       then push xys x
            drop xs
            x' <- pull xs
            jump loop x' y
       else push xys y
            drop ys
            y' <- pull ys
            jump loop x y'
  in x <- pull xs
     y <- pull ys
     jump loop x y
\end{code}

@@

@@

\section{The problem with synchronised product}
Synchronised product is: both machines can take independent steps so long as it is not in the alphabet of the other, but have to agree on any shared actions.
So if both machines pull from @xs@, the @pull xs@ message can only be executed when both machines agree.

@@

\begin{code}
dup ins = process l r.
  let loop =
     i <- pull ins
     push l i
     push r i
     drop ins
     jump loop
  in jump loop
\end{code}
And the modified zip uses this like so:

>>>>

Furthermore, existing systems for combinator fusion only support a subset of combinators, and only in limited cases.
Pull fusion supports joins where one combinator pulls from multiple inputs, but not splits where one output is used my multiple combinators.
Pull can thus express @merge@, but not @partition@.
Push fusion supports splits (@partition@), but not joins (@merge@).
Polarised fusion~\cite{lippmeier2016polarized} addresses these shortcomings by separating the computation into pull producers and push consumers, with explicit drain loops to convert between the two sections. However this requires manual plumbing.

@@

\section{Streams vs Push/Pull}

Types
\begin{code}
Stream  : Data -> Static
Source  : Data -> Static
Sink    : Data -> Static
Process :         Static
\end{code}

Streams have the advantage of being able to be passed between producer and consumer.
Stream combinators are easily composed as normal functions.

The polar representation of Source/Sink has the advantage of having obvious bufferless execution semantics.
Composition is harder.

\section{Processes}
\label{s:Process}
We will start by looking at a combinator that increments each value.
Given an input stream @ins@, @increment ins@ returns a new stream with the incremented values.
The function @stream_1_1@ takes a process program with one pull input and one push output, and converts it to a stream program.

The process blocks reading the value from @ins@.
After reading the value @i@, it pushes @i + 1@ to the @out@ stream.
The pulled value must be dropped or released before another value can be performed.
Operationally this is a no-op, but is used for coordinating between producer and consumer.
Finally, the process loops back to the start.

\begin{code}
increment : Stream Int -> Stream Int
increment = stream_1_1 \ins out.
  letrec
    p1   = pull ins       p2
    p2 i = push out (i+1) p3
    p3   = drop ins       p1
  in p1
\end{code}

@@

\begin{code}
stream_1_1 : (Source a -> Sink b
              -> Process)
          -> Stream a -> Stream b
stream_2_1 : (Source a -> Source b -> Sink c
              -> Process)
          -> Stream a -> Stream b -> Stream c
stream_1_2 : (Source a -> Sink b -> Sink c
              -> Process)
          -> Stream a -> (Stream b, Stream c)

pull : Source a -> (a -> Process) -> Process
push : Sink   a ->  a -> Process  -> Process
drop : Source a ->       Process  -> Process
case : Bool -> Process -> Process -> Process
\end{code}

@@

\begin{code}
map f = stream_1_1 \is os.
  letrec
    p1   = pull is p2
    p2 i = push os (f i) p3
    p3   = drop is p1
  in p1
\end{code}

\begin{code}
filter f = stream_1_1 \is os.
  letrec
    p1   = pull is p2
    p2 i = case (f x) (p3 i) p4
    p3 i = push os i p4
    p4   = drop is p1
  in p1
\end{code}

\begin{code}
partition f = stream_1_2 \is ots ofs.
  letrec
    p1   = pull is p2
    p2 i = case (f x)
            (push ts i p3)
            (push fs i p3)
    p3   = drop is p1
  in p1
\end{code}

\begin{code}
zip = stream_2_1 \xs ys xys.
  letrec
    p1     = pull xs        p2
    p2 x   = pull ys        p3
    p3 x y = push xys (x,y) p4
    p4     = drop xs        p5
    p5     = drop ys        p1
  in p1
\end{code}

\begin{code}
merge = stream_2_1 \xs ys xys.
  letrec
    go x y = case (x < y)
             (pX x y)
             (pY x y)
    pX x y = push xys x
            (drop xs
            (pull xs (\x'. go x' y)))
    pY x y = push xys y
            (drop ys
            (pull ys (\y'. go x y')))
  in pull xs (\x. pull ys (\y. go x y))
\end{code}

@@

\begin{code}
channels : Process -> ({Source}, {Sink})
channels (pull i   p) = ({i}, {})  U channels p
channels (drop i   p) = ({i}, {})  U channels p
channels (push o _ p) = ({}, {o})  U channels p
channels (case _ p q) = channels p U channels q
\end{code}

@@

\section{The problem with synchronised product}
Related work on Network Fusion~\cite{fradet2004network} allows fusion of Kahn Process Networks using synchronised product to fuse pairs of processes together.
Synchronised product is: both machines can take independent steps so long as it is not in the alphabet of the other, but have to agree on any shared actions.
So if both machines pull from @xs@, the @pull xs@ message can only be executed when both machines agree.

@@

\begin{code}
dup = stream_1_2 \is ols ors.
  letrec
    p1   = pull is p2
    p2 i = push ols i (p3 i)
    p3 i = push ors i p4
    p4   = drop is p1
  in p1
\end{code}
And the modified zip uses this like so:
